Longitudinal Data Cleaning - A Case of the NLSY79 Data

### 2021-02-23 report/report.Rmd
Add plots to IDA section.

As stated previously, the respondents' ages ranged from 12 to 22 when first interviewed in 1979. Hence, we would like to validate whether all of the respondents were in this range. For the demographic data validation, we used the `full_demographic` i.e. the data with the survey years 1979 sample.

```{r valid-age, echo = FALSE, fig.cap="Subjects' Age Distribution"}
age <- yowie::demographic_nlsy79 %>%
  ggplot(aes(x = age_1979)) +
  geom_bar(fill = "#45B39D")

gender <- yowie::demographic_nlsy79 %>%
  ggplot(aes(x = gender,
             fill = gender)) +
  geom_bar() +
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, colour = "white")+
  scale_fill_manual("gender", values = c("MALE" = "#2471A3", "FEMALE" = "#9B59B6"))+
  theme(legend.position = "none")

race <- yowie::demographic_nlsy79 %>%
  ggplot(aes(x = race,
             fill = race)) +
  geom_bar() +
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, colour = "white") +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 10))

@@

>>>>

As stated previously, the respondents' ages ranged from 12 to 22 when first interviewed in 1979. Hence, we would like to validate whether all of the respondents were in this range. Additionally, the [NLSY](https://www.nlsinfo.org/content/cohorts/nlsy79/intro-to-the-sample/nlsy79-sample-introduction) also provided the number of the survey cohort by their gender (6,403 males and 6,283 females) and race (7,510 Non-Black/Non-Hispanic; 3,174 Black; 2,002 Hispanic). To validate this, we used the `full_demographic` i.e. the data with the survey years 1979 sample. Figure \ref{fig:demoprofile} suggests that the demographic data we had is consistent with the sample information in the database.

```{r demoprofile, echo = FALSE, fig.cap="Demography Profile of the NLSY79 Cohort", fig.width=6}
age <- yowie::demographic_nlsy79 %>%
  ggplot(aes(x = age_1979)) +
  geom_bar(fill = "#45B39D") +
  ggtitle("age distribution") +
  theme(plot.title = element_text(size = 10))

gender <- yowie::demographic_nlsy79 %>%
  ggplot(aes(x = gender,
             fill = gender)) +
  geom_bar() +
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, colour = "white")+
  scale_fill_manual("gender", values = c("MALE" = "#2471A3", "FEMALE" = "#9B59B6"))+
  theme(legend.position = "none")  +
  ggtitle("gender_distribution") +
  theme(plot.title = element_text(size = 10))

race <- yowie::demographic_nlsy79 %>%
  ggplot(aes(x = race,
             fill = race)) +
  geom_bar() +
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, colour = "white") +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 10)) +
  ggtitle("race_distribution") +
  theme(plot.title = element_text(size = 10),
        axis.text.x = element_text(size = 6))

@@

The next step is that we explored the mean hourly wage data, in this case, we only explored the wages data in `wages_demog_hs`. Figure \ref{fig:featureplot} conveys that there is clearly a problem in the mean hourly wage values. Figure \ref{fig:featureplot} a.) shows that some observations had an exceptionally high figure of wages, even more than US$10,000 per hour. In Figure \ref{fig:featureplot} b.), we barely see any difference in the minimum, median, an maximum value of the wages since the distribution is heavily skewed to the right. Additionally, Table \ref{tab:summarytable} shows that the overall wages median of the cohort is only 7.2, while the mean is 11.87. It indicates that the data might contains a lot of extreme values.

```{r spaghetti, echo = FALSE}
spag <- wages_demog_hs %>%
  ggplot(aes(x = year,
             y = mean_hourly_wage,
             group = id)) +
  geom_line(alpha = 0.1) +
  ggtitle("a)") +
  theme(plot.title = element_text(size = 10))
```

```{r featureplot, echo = FALSE, fig.cap = "a. The mean hourly wage of cohort whose highest grade completed is up to 12th grade, b. Three features of the cohort", fig.width=6, fig.height=4}
wages_hs2020 <- as_tsibble(x = wages_demog_hs,
                    key = id,
                    index = year,
                    regular = FALSE)
wages_three_feat <- wages_hs2020 %>%
  features(mean_hourly_wage, 
           feat_three_num
           )
wages_feat_long <- wages_three_feat %>%
  pivot_longer(c(min, med, max), names_to = "feature", values_to = "value")
 feature <- ggplot(wages_feat_long) +
  geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
   ggtitle("b)") +
   theme(plot.title = element_text(size = 10))

spag + feature
```
```{r summarytable, echo = FALSE}
kable(as.array(summary(wages_demog_hs$mean_hourly_wage)), 
      caption = "Summary Statistics of Wages of High School Data",
      col.names = c("Statistics", "Value")) %>%
  kable_styling(latex_options = "striped")
```

Further, we took 36 samples randomly from the data and plotted it as is seen in Figure \ref{fig:sampleplot}. It implies that not only that some observations earned an extremely high figures of wages, but some also had a very fluctuative wages, fo example in panel number 5, 7, and 11.

```{r sampleplot, echo = FALSE, fig.cap="Some samples of observations"}
set.seed(20210225)

ggplot(wages_hs2020, 
       aes(x = year,
                y = mean_hourly_wage,
                group = id)) +
  geom_line(alpha = 0.7) +
  facet_sample() +
  ylab("mean hourly wage") 
```

### 2021-02-23 report/report.tex
Add plots to IDA section.

As stated previously, the respondents' ages ranged from 12 to 22 when
first interviewed in 1979. Hence, we would like to validate whether all
of the respondents were in this range. For the demographic data
validation, we used the \texttt{full\_demographic} i.e.~the data with
the survey years 1979 sample.

\begin{figure}
\centering
\includegraphics{report_files/figure-latex/valid-age-1.pdf}
\caption{Subjects' Age Distribution}
\end{figure}

>>>>

As stated previously, the respondents' ages ranged from 12 to 22 when
first interviewed in 1979. Hence, we would like to validate whether all
of the respondents were in this range. Additionally, the
\href{https://www.nlsinfo.org/content/cohorts/nlsy79/intro-to-the-sample/nlsy79-sample-introduction}{NLSY}
also provided the number of the survey cohort by their gender (6,403
males and 6,283 females) and race (7,510 Non-Black/Non-Hispanic; 3,174
Black; 2,002 Hispanic). To validate this, we used the
\texttt{full\_demographic} i.e.~the data with the survey years 1979
sample. Figure \ref{fig:demoprofile} suggests that the demographic data
we had is consistent with the sample information in the database.

\begin{figure}
\centering
\includegraphics{report_files/figure-latex/demoprofile-1.pdf}
\caption{Demography Profile of the NLSY79 Cohort}
\end{figure}

The next step is that we explored the mean hourly wage data, in this
case, we only explored the wages data in \texttt{wages\_demog\_hs}.
Figure \ref{fig:featureplot} conveys that there is clearly a problem in
the mean hourly wage values. Figure \ref{fig:featureplot} a.) shows that
some observations had an exceptionally high figure of wages, even more
than US\$10,000 per hour. In Figure \ref{fig:featureplot} b.), we barely
see any difference in the minimum, median, an maximum value of the wages
since the distribution is heavily skewed to the right. Additionally,
Table \ref{tab:summarytable} shows that the overall wages median of the
cohort is only 7.2, while the mean is 11.87. It indicates that the data
might contains a lot of extreme values.

\includegraphics{report_files/figure-latex/featureplot-1.pdf}

\begin{table}

\caption{\label{tab:summarytable}Summary Statistics of Wages of High School Data}
\centering
\begin{tabular}[t]{l|r}
\hline
Statistics & Value\\
\hline
\rowcolor{gray!6}  Min. & 0.01000\\
\hline
1st Qu. & 4.50000\\
\hline
\rowcolor{gray!6}  Median & 7.20000\\
\hline
Mean & 11.86578\\
\hline
\rowcolor{gray!6}  3rd Qu. & 11.74000\\
\hline
Max. & 33400.00000\\
\hline
\end{tabular}
\end{table}

Further, we took 36 samples randomly from the data and plotted it as is
seen in Figure \ref{fig:sampleplot}. It implies that not only that some
observations earned an extremely high figures of wages, but some also
had a very fluctuative wages, fo example in panel number 5, 7, and 11.

\begin{figure}
\centering
\includegraphics{report_files/figure-latex/sampleplot-1.pdf}
\caption{Some samples of observations}
\end{figure}
